---
title: 'Texas: Test of Transformations'
author: "Xianbin Cheng"
date: "March 15, 2019"
output: html_document
---

# Objective
  
  * Convert AF_class and FM_class to a binary vector
  * Try log(1/R) transformation to X
  * Try logR transformation to X
  * Try Savitzky-Golay smooth filtering to each row of X
  
# Method

1. Load libraries and files. Convert class vectors to binary vectors.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(signal)
```

```{r}
X_raw = read.csv("TX X_Conc_1680_obs.csv", header = TRUE, row.names = 1, stringsAsFactors = FALSE)
X_raw$FM_class = gsub(pattern = "M", replacement = "H", x = X_raw$FM_class) %>% as.factor()
X_raw$AF_class = as.factor(X_raw$AF_class)
summary(X_raw$AF_class)
summary(X_raw$FM_class)
```

2. Load necessary functions.

```{r}
## Convert a wide dataset into a long dataset
my_melt = function(df){
  df1 = df %>%
    gather(., key = "nm", value = "value", -c(Kernel_ID, Spec_ID_all, AF_class, FM_class))
  df1$nm = gsub(pattern = "X", replacement = "", x = df1$nm) %>% as.numeric()
  return(df1)
}

## Extract the observation at the 5th percentile and 95th percentile
lb <- function(x, probs = 0.05){quantile(x, probs)}
ub <- function(x, probs = 0.95){quantile(x, probs)}

## Generate spectral graph of the medians at each wavelength with 5th - 95th percentile interval
spec_plot = function(data, class, yaxis = "Intensity", legend = "5th - 95th Percentile"){
  
  ### Generate columns of observations at 5th, 50th and 95th percentile
  df = data %>%
    group_by_(class, "nm") %>%
    summarise(lower = lb(value), med = median(value), upper = ub(value))
  
  ### Visualize
  ggplot(data = df) +
    geom_line(aes_string(x = "nm", y = "med", color = class)) +
    geom_ribbon(aes_string(x = "nm", ymin = "lower", ymax = "upper", fill = class), alpha = 0.3) +
    labs(x = "Wavelength (nm)", y = yaxis, color = "Median", fill = legend) +
    scale_color_discrete(labels = c("High", "Low", "Medium")) +
    scale_fill_discrete(labels = c("High", "Low", "Medium")) +
    theme_classic() +
    theme(legend.position = "top")
}
```

3. Convert wide data to long data.

```{r}
X_raw_long = my_melt(df = X_raw)
```

4. Transformations.

```{r}
# First shift the spectra upward by 20 so that all numbers are positive
min(X_raw_long$value)

X_raw_long_pos = X_raw_long
X_raw_long_pos$value = X_raw_long_pos$value + 20

# log(R) transformation
X_raw_long_logR = X_raw_long_pos
X_raw_long_logR$value = log(X_raw_long_logR$value)

# log(1/R) = absorbance
X_raw_long_abs = X_raw_long_pos
X_raw_long_abs$value = log(1/X_raw_long_logR$value)

# Savitsky-Golay Filtering
X = X_raw[,5:ncol(X_raw)] %>%
  as.matrix() %>%
  t()
X_sg = apply(X = X, MARGIN = 2, FUN = sgolayfilt, p = 3) %>%
  t() %>%
  cbind(X_raw[,1:4], .)
colnames(X_sg) = colnames(X_raw)

X_sg_long = my_melt(df = X_sg)
```

5. Produce plots.

```{r}
## Aflatoxin: Raw - Background
af_raw_plot = spec_plot(data = X_raw_long, class = "AF_class", yaxis = "Intensity (Raw - Background)")
fm_raw_plot = spec_plot(data = X_raw_long, class = "FM_class", yaxis = "Intensity (Raw - Background)")
```

```{r}
# Log(R)
af_logR_plot = spec_plot(data = X_raw_long_logR, class = "AF_class", yaxis = "log Intensity (upshifted by 20)")
fm_logR_plot = spec_plot(data = X_raw_long_logR, class = "FM_class", yaxis = "log Intensity (upshifted by 20)")
```

```{r}
# log(1/R)
af_abs_plot = spec_plot(data = X_raw_long_abs, class = "AF_class", yaxis = "Absorbance")
fm_abs_plot = spec_plot(data = X_raw_long_abs, class = "FM_class", yaxis = "Absorbance")
```

```{r}
#Savitzky-Golay Filter
af_sg_plot = spec_plot(data = X_sg_long, class = "AF_class", yaxis = "Intensity (Raw - Background)")
fm_sg_plot = spec_plot(data = X_sg_long, class = "FM_class", yaxis = "Intensity (Raw - Background)")
```


# Result

1. Visualization of raw spectra. Left: Aflatoxin. Right: Fumonisin.

```{r, echo = FALSE, fig.show = "hold", out.width = "50%"}
af_raw_plot
fm_raw_plot
```

2. Visualization of spectra transformed by log(R). 

```{r, echo = FALSE, fig.show = "hold", out.width = "50%"}
af_logR_plot
fm_logR_plot
```

3. Visualization of spectra transformed by log(1/R)

```{r, echo = FALSE, fig.show = "hold", out.width = "50%"}
af_abs_plot
fm_abs_plot
```

4. Visualization of spectra pre-processed with Savitzky-Golay smoothing filter.

```{r, echo = FALSE, fig.show = "hold", out.width = "50%"}
af_sg_plot
fm_sg_plot
```


