---
title: "Texas Spectra Cleaning"
author: "Xianbin Cheng"
date: "March 7, 2019"
output: html_document
---

# Objective

  * Compare the backgrounds and references during the Texas corn 2017 period
  * Re-make normalized spectra data from the original design matrix X
  * Make percent reflectance spectra data from 

# Method

1. Load libraries and files.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)
```

```{r}
X = read.csv("TX X_Conc_1680_obs.csv", header = TRUE, row.names = 1)
X$AF_class = factor(X$AF_class, levels = c("H", "M", "L"))
X$FM_class = factor(X$FM_class, levels = c("H", "M", "L"))

str(X, list.len = 10)
```

```{r}
# Backgrounds and references
Bg_Rf = read.csv("TX Bg_Rf.csv", header = TRUE, stringsAsFactors = FALSE)
str(Bg_Rf, list.len = 10)
```

2. Re-make the normalized data by centering and scaling the `X`.

```{r, eval = FALSE}
temp = X[ ,5:ncol(X)] %>%
  scale(x = ., center = TRUE, scale = TRUE)

X_norm = cbind(X[ ,1:4], temp)
```

3. Clean up the spectra transformed by percent reflectance. The remaining merging with concentration information is conducted in 'Texas Data Cleaning 6.Rmd'.

```{r, eval = FALSE}
# Percent reflectance
perc1 = dir(pattern = "X1_perc")
perc2 = dir(pattern = "X2_perc")

perc1_csv = map(.x = perc1, .f = read.csv, header = TRUE, stringsAsFactors = FALSE) %>%
  bind_rows()

perc2_csv = map(.x = perc2, .f = read.csv, header = TRUE, stringsAsFactors = FALSE) %>%
  bind_rows()
```

```{r, echo = FALSE, eval = FALSE}
write.csv(x = perc1_csv, file = "TX X1_perc.csv", row.names = TRUE)
write.csv(x = perc2_csv, file = "TX X2_perc.csv", row.names = TRUE)
```

4. Clean the backgrounds and references.

```{r, eval = FALSE}
batches = list(t09_20 = c("HH1", "LL1"),
                     t09_22 = c("HH1D"),
                     t10_18 = c("LH1", "LH2", "MH1", "MH2"),
                     t10_19 = c("LH1D","LH2D", "MH1D", "MH2D"),
                     t11_02 = c("HH2", "LL2", "HL1", "HL2", "LH3", "LH4", "ML1", "ML2"),
                     t11_07 = c("HH2D", "LL2D", "LH3D", "LH4D", "HL1D", "HL2D", "ML1D", "ML2D"),
                     t11_14 = c("LM1", "LM2", "MM1", "MM2", "HM1", "HM2"),
                     t11_15 = c("LM1D",  "LM2D", "MM1D", "MM2D", "HM1D", "HM2D")
                     )

spec_meta = data.frame(Date = names(unlist(batches)),
                      Spec = unlist(batches), row.names = NULL) %>%
  mutate(Date = substr(Date, start = 2, stop = 6),
         Bulk_class = substr(Spec, start = 1, stop = 2),
         Plate = substr(Spec, start = 3, stop = 3),
         Duplicate = str_detect(string = Spec, pattern = "D"))
```

```{r, echo = FALSE}
spec_meta = read.csv("TX spectra metadata.csv", header = TRUE, stringsAsFactors = FALSE, row.names = 1)
str(spec_meta)
```

5. Visualize backgrounds and references.

```{r}
## Convert a wide dataset into a long dataset
my_melt = function(df){
  df1 = df %>%
    gather(., key = "nm", value = "value", -c(Date, Type, Name))
  df1$nm = gsub(pattern = "X", replacement = "", x = df1$nm) %>% as.numeric()
  return(df1)
}

Bg_Rf_long = my_melt(df = Bg_Rf)
```

# Result

1. Visualize the backgrounds.

```{r, fig.width = 13}
ggplot(data = subset(x = Bg_Rf_long, subset = Type == "Bg")) +
  geom_line(aes_string(x = "nm", y = "value", color = "Date")) +
  labs(x = "Wavelength (nm)", y = "Intensity", color = "Date") +
  theme_bw() +
  theme(legend.position = "right")
```

2. Visualize the references (corrected by Raw A/D counts - Background A/D counts).

```{r, fig.width = 13}
ggplot(data = subset(x = Bg_Rf_long, subset = Type == "Rf")) +
  geom_line(aes_string(x = "nm", y = "value", color = "Date")) +
  labs(x = "Wavelength (nm)", y = "Intensity", color = "Date") +
  theme_bw() +
  theme(legend.position = "right")
```

3. A summary table of spectra metadata.

```{r}
kable_styling(kable(spec_meta, format = "html"), full_width = FALSE)
```

