---
title: "Texas Data Pre-processsing"
author: "Xianbin Cheng"
date: "March 6, 2019"
output: html_document
---

# Objective

  * Pre-process the spectral data with PCA
  
# Method

1. Load libraries and files.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)
```

```{r}
spec_meta = read.csv("TX Spectra Metadata.csv", header = TRUE, row.names = 1, stringsAsFactors = FALSE)
visual_meta = read.csv("TX Visual Metadata.csv", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

X = read.csv("TX X_Conc_1680_obs.csv", header = TRUE, row.names = 1)
X$AF_class = factor(X$AF_class, levels = c("H", "M", "L"))
X$FM_class = factor(X$FM_class, levels = c("H", "M", "L"))

X_norm = read.csv("TX X_norm_Conc_1680_obs.csv", header = TRUE, row.names = 1)
X_norm$AF_class = factor(X_norm$AF_class, levels = c("H", "M", "L"))
X_norm$FM_class = factor(X_norm$FM_class, levels = c("H", "M", "L"))

X_perc = read.csv("TX X_perc_Conc_1680_obs.csv", header = TRUE, row.names = 1)
X_norm$AF_class = factor(X_norm$AF_class, levels = c("H", "M", "L"))
X_norm$FM_class = factor(X_norm$FM_class, levels = c("H", "M", "L"))
```

2. Transform `X` and `X_norm` by PCA. Draw 2D PCA plots. 

```{r}
pca_X = prcomp(x = X[ ,5:ncol(X)], center = FALSE, scale. = FALSE)
pca_X_norm = prcomp(x = X_norm[ ,5:ncol(X_norm)], center = FALSE, scale. = FALSE)
pca_X_perc = prcomp(x = X_perc[ ,5:ncol(X_perc)], center = FALSE, scale. = FALSE)
```

3. Extract the metadata.

```{r}
Meta = X_norm %>%
  dplyr::select(Kernel_ID:Spec_ID_all) %>%
  mutate(Spec = str_split(string = Spec_ID_all, pattern = "-", simplify = TRUE)[,1]) %>%
  left_join(x = ., y = spec_meta, by = "Spec") %>%
  dplyr::select(-c(Plate, Duplicate)) %>%
  left_join(x = ., y = visual_meta, by = "Kernel_ID")
```

3. Based on the PCA plots, draw a dividing line and identify the outliers.

```{r}
# Find out the outliers PC2 > PC1 * (25/-75) + 25
out = function(data, k = k, b = b){
  if(data[2] < data[1] * k + b){
    return(TRUE)
  } else {
    return(FALSE)
  }
}

# Find the points that fall beyond PC2 = PC1 * k + b
k = 0.5
b = -75/2

test = pca_X_norm$x[,1:2] 
test2 = apply(X = test, MARGIN = 1, FUN = out, k = k, b = b)

# Find the rows of outliers
idx = test2[test2 == TRUE] %>% 
  names() %>%
  as.numeric()

# Find the outliers
outliers = Meta[idx, ] 

# Create an outlier factor vector
outlier_binary = vector(mode = "numeric", length = nrow(X_norm)) + 0
outlier_binary[idx] = 1
outlier_binary = as.factor(outlier_binary)
summary(outlier_binary)
```

3. Find out about the correlation between outliers and mycotoxin levels, date when we collected spectra, BGYF, discoloration, brokenness, sample source.

```{r}
# mycotoxins
tab_AF = table(outlier_binary, X_norm$AF_class)
tab_FM = table(outlier_binary, X_norm$FM_class)

# date
tab_date = table(Meta$Date, outlier_binary)

# BGYF
tab_bgyf = table(Meta$BGYF, outlier_binary)

# discoloration
tab_discoloration = table(Meta$Discoloration, outlier_binary)

# brokenness
tab_broken = table(Meta$Brokenness, outlier_binary)

# sample source
tab_ss = table(Meta$Bulk_class, outlier_binary)
```

# Result

1. PCA plots.

```{r, echo = FALSE}
draw_pca_2d = function(data, class, divide, slope, intercept){
  
  a = data$x %>%
    as.data.frame()
  b = cbind(class, a)
  
  c = ggplot(data = b, aes(x = PC1, y = PC2)) +
    geom_point(aes(color = class)) +
    theme_bw()
  
  if(divide == TRUE){
    d = c +
      geom_abline(slope = slope, intercept = intercept)
    
    return(d)
  } else {
    return(c)
  }
}
```

```{r, fig.show="hold", out.width= "33%"}
# Aflatoxin
# Untransformed X
draw_pca_2d(data = pca_X, class = X$AF_class, divide = FALSE)

# Normalized X_norm
draw_pca_2d(data = pca_X_norm, class = X$AF_class, divide = FALSE)

# Percent reflectance
draw_pca_2d(data = pca_X_perc, class = X$AF_class, divide = FALSE)
```

```{r, fig.show= "hold", out.width= "33%"}
# Fumonisin
# Untransformed X
draw_pca_2d(data = pca_X, class = X$FM_class, divide = FALSE)

# Normalized X_norm
draw_pca_2d(data = pca_X_norm, class = X$FM_class, divide = FALSE)

# Percent refletance X_perc
draw_pca_2d(data = pca_X_perc, class = X$FM_class, divide = FALSE)
```

2. Correlation

```{r, out.size= "50%"}
# AF
draw_pca_2d(data = pca_X_norm, class = X_norm$AF_class, divide = TRUE, slope = k, intercept = b)
tab_AF
fisher.test(x = tab_AF)

# FM
draw_pca_2d(data = pca_X_norm, class = X_norm$FM_class, divide = TRUE, slope = k, intercept = b)
tab_FM
fisher.test(x = tab_FM, hybrid = TRUE)

# Spectra collection date
draw_pca_2d(data = pca_X_norm, class = Meta$Date, divide = TRUE, slope = k, intercept = b)
tab_date
fisher.test(x = tab_date, hybrid = TRUE)

# BGYF
draw_pca_2d(data = pca_X_norm, class = as.factor(Meta$BGYF), divide = TRUE, slope = k, intercept = b)
tab_bgyf
fisher.test(x = tab_bgyf)

# Brokenness
draw_pca_2d(data = pca_X_norm, class = as.factor(Meta$Brokenness), divide = TRUE, slope = k, intercept = b)
tab_broken
fisher.test(x = tab_broken)

# discoloration
draw_pca_2d(data = pca_X_norm, class = as.factor(Meta$Discoloration), divide = TRUE, slope = k, intercept = b)
tab_discoloration
fisher.test(x = tab_discoloration)

# Sample source (Bulk-class)
draw_pca_2d(data = pca_X_norm, class = Meta$Bulk_class, divide = TRUE, slope = k, intercept = b)
tab_ss
fisher.test(x = tab_ss, hybrid = TRUE)
```

# Conclusion

  * The Fisher's exact test shows that bulk class (i.e. sample source) and brokenness are associated with whether a kernel is an outlier or not. However, it seems that when the kernel is NOT an outlier, it's more likely to be broken. And we see an association between outliers and bulk class simply because there are much more points in the left of the dividing line.
  
  
