---
title: 'Texas: Data Cleaning 4'
author: "Xianbin Cheng"
date: "February 4, 2019"
output: html_document
---

# Objective

  * Clean up the NA's in the Texas ELISA data before 04/12/2018
  
# Method

1. Load in necessary libraries and files.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
```

2. Summary before cleaning.

```{r}
conc = read.csv("Texas Conc and Class 04_12_2018.csv", header = TRUE, stringsAsFactors = FALSE, row.names = 1)
conc_preclean = conc %>%
  mutate(plate = str_split(string = conc$Kernel_ID, pattern = "-", simplify = TRUE)[,1],
         kernel_ID = str_split(string = conc$Kernel_ID, pattern = "-", simplify = TRUE)[,2] %>% as.numeric()) %>%
  arrange(.data = ., plate, kernel_ID) %>%
  group_by(plate)

str(conc_preclean)
```

2. Load in the cleaned version.

```{r}
conc2 = read.csv("Texas Conc and Class 02_05_19.csv", header = TRUE, stringsAsFactors = FALSE, row.names = 1)
str(conc2)
```

# Result

1. Summary before cleaning.

```{r}
# Aflatoxin
summary(conc$AF_class %>% as.factor())
# Fumonisin
summary(conc$FM_class %>% as.factor())

# How many plates? How many kernels per plate?
conc_preclean %>%
  summarise(obs = n())

##### How many NAs?
count_na = function(data, type){
  if(type == "AF"){
    a = "AF.ppb."
  } else if (type == "FM"){
    a = "FM.ppm."
  }
  sum(is.na(data[[a]]))
}

# How many NA's in aflatoxin?
conc_preclean %>%
  split(x = ., f = conc_preclean$plate) %>%
  map(.x = ., .f = count_na, type = "AF") %>%
  unlist()

# How many Na's in fumonisin?
conc_preclean %>%
  split(x = ., f = conc_preclean$plate) %>%
  map(.x = ., .f = count_na, type = "FM") %>%
  unlist()
```

2. Summary after cleaning.

```{r}
# How many AFs and FMs?
AF_class2 = ifelse(test = conc2$AF.ppb. < 20, yes = "L", no = ifelse(test = conc2$AF.ppb. >= 50, yes = "H", no = "M")) %>% 
  as.factor()
summary(AF_class2)

FM_class2 = ifelse(test = conc2$FM.ppm. < 1, yes = "L", no = ifelse(test = conc2$FM.ppm. >= 4, yes = "H", no = "M")) %>% 
  as.factor()
summary(FM_class2)
```

```{r}
# Check for NA's
anyNA(conc2)
```

```{r}
# Check the mediums for FM. Find out which kernel is the extra medium after cleaning.

## Before cleaning
subset(x = conc, subset = FM_class == "M") %>%
  select(Kernel_ID, FM.ppm.)

## After cleaning
subset(x = conc2, subset = FM_class2 == "M") %>%
  select(Kernel_ID, FM.ppm.)
```

# Conclusion

    There are 528 kernels involved in this part of the experiment. There is 1 more kernel categorized as "medium in fumonisin" after cleaning. It is confirmed by checking the original ELISA data that "201-16" has the correct fumonisin concentration after cleaning.
    
```{r, echo = FALSE, eval = FALSE}
temp = conc2 %>%
  mutate(AF_class = AF_class2,
         FM_class = FM_class2)

colnames(temp)[c(6,7)] = c("<LOD.AF", "<LOD.FM")
write.csv(x = temp, file = "Texas_Conc_Class_528obs_02_05_19.csv")
```

