---
title: "Texas Visualization 6"
author: "Xianbin Cheng"
date: "March 6, 2019"
output: html_document
---

# Objective

  * Visualize spectra of Texas corn 2017 (1680 spectra)

## Method ##

1. Load the libraries and read in the files.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(reshape2)
```

```{r}
X_raw = read.csv("TX X_Conc_1680_obs.csv", header = TRUE, row.names = 1)
X_norm = read.csv("TX X_norm_Conc_1680_obs.csv", header = TRUE, row.names = 1)
X_perc = read.csv("TX X_perc_Conc_1680_obs.csv", header = TRUE, row.names = 1)

str(X_raw, list.len = 10)
```

```{r}
summary(X_raw$AF_class)
summary(X_raw$FM_class)
```

2. Load the custom functions.

```{r}
## Convert a wide dataset into a long dataset
my_melt = function(df){
  df1 = df %>%
    gather(., key = "nm", value = "value", -c(Kernel_ID, Spec_ID_all, AF_class, FM_class))
  df1$nm = gsub(pattern = "X", replacement = "", x = df1$nm) %>% as.numeric()
  return(df1)
}

## Extract the observation at the 5th percentile and 95th percentile
lb <- function(x, probs = 0.05){quantile(x, probs)}
ub <- function(x, probs = 0.95){quantile(x, probs)}

## Generate spectral graph of the medians at each wavelength with 5th - 95th percentile interval
spec_plot = function(data, class, yaxis = "Intensity", legend = "5th - 95th Percentile"){
  
  ### Generate columns of observations at 5th, 50th and 95th percentile
  df = data %>%
    group_by_(class, "nm") %>%
    summarise(lower = lb(value), med = median(value), upper = ub(value))
  
  ### Visualize
  ggplot(data = df) +
    geom_line(aes_string(x = "nm", y = "med", color = class)) +
    geom_ribbon(aes_string(x = "nm", ymin = "lower", ymax = "upper", fill = class), alpha = 0.3) +
    labs(x = "Wavelength (nm)", y = yaxis, color = "Median", fill = legend) +
    scale_color_discrete(labels = c("High", "Low", "Medium")) +
    scale_fill_discrete(labels = c("High", "Low", "Medium")) +
    theme_classic() +
    theme(legend.position = "top")
}
```

3. Convert all datasets into long datasets.

```{r}
X_raw_long = my_melt(df = X_raw)
X_norm_long = my_melt(df = X_norm)
X_perc_long = my_melt(df = X_perc)
```

4. Generate plots.

```{r}
## Aflatoxin: Raw - Background
af_raw_plot = spec_plot(data = X_raw_long, class = "AF_class", yaxis = "Intensity (Raw - Background)")

## Aflatoxin: Normalized
af_norm_plot = spec_plot(data = X_norm_long, class = "AF_class", yaxis = "Intensity (Normalized)")

## AF: percent reflectance
af_perc_plot = spec_plot(data = X_perc_long, class = "AF_class", yaxis = "Intensity (Percent Reflectance)")

## Fumonisin: Raw - Background
fm_raw_plot = spec_plot(data = X_raw_long, class = "FM_class", yaxis = "Intensity (Raw - Background)")

## Fumonisin: Normalized
fm_norm_plot = spec_plot(data = X_norm_long, class = "FM_class", yaxis = "Intensity (Normalized)")

## FM: percent reflectance
fm_perc_plot = spec_plot(data = X_perc_long, class = "FM_class", yaxis = "Intensity (Percent Reflectance)")
```

5. Extra plots: convert three-class labels to binary labels by treating mediums as highs as well. Since only fumonisin has "M"s, we are only gonna convert `FM_class`.

```{r}
# Untranformed
X_raw_bi = X_raw
X_raw_bi$FM_class = gsub(pattern = "M", replacement = "H", x = X_raw_bi$FM_class) %>%
  as.factor()
summary(X_raw_bi$FM_class)
X_raw_long_bi = my_melt(df = X_raw_bi)
fm_raw_bi_plot = spec_plot(data = X_raw_long_bi, class = "FM_class", yaxis = "Intensity (Raw - Background)")

# Normalized
X_norm_bi = X_norm
X_norm_bi$FM_class = gsub(pattern = "M", replacement = "H", x = X_norm_bi$FM_class) %>%
  as.factor()
summary(X_norm_bi$FM_class)
X_norm_long_bi = my_melt(df = X_norm_bi)
fm_norm_bi_plot = spec_plot(data = X_norm_long_bi, class = "FM_class", yaxis = "Intensity (Normalized)")

# Fumonisin: Percent Reflectance
X_perc_bi = X_perc
X_perc_bi$FM_class = gsub(pattern = "M", replacement = "H", x = X_perc_bi$FM_class) %>%
  as.factor()
summary(X_perc_bi$FM_class)
X_perc_long_bi = my_melt(df = X_perc_bi)
fm_perc_bi_plot = spec_plot(data = X_perc_long_bi, class = "FM_class", yaxis = "Intensity (Percent Reflectance)")
```

## Result ##

1. Spectral plot for aflatoxin

```{r, echo = FALSE}
af_raw_plot
af_norm_plot
af_perc_plot
```

2. Spectral plot for fumonisin.

```{r, echo = FALSE}
fm_raw_plot
fm_norm_plot
fm_perc_plot
```

```{r}
# Binary class 
fm_raw_bi_plot
fm_norm_bi_plot
fm_perc_bi_plot
```

```{r, echo = FALSE, eval = FALSE}
pdf(file = "Texas Corn 2017 Spectra.pdf")
  af_raw_plot
  af_norm_plot
  af_perc_plot
  fm_raw_plot
  fm_norm_plot
  fm_perc_plot
  fm_raw_bi_plot
  fm_norm_bi_plot
  fm_perc_bi_plot
dev.off()
```
