---
title: "Texas Corn ELISA Visualization"
author: "Xianbin Cheng"
date: "April 12, 2018"
output: html_document
---

## Method ##

### Data Cleaning ###

1. Load the libraries and read in the file.
```{r, message = FALSE, warning = FALSE}
library(tidyverse)

df = read.csv("Texas Corn Total ELISA.csv", header = TRUE)
```

2. Only keep the following Kernel_IDs: 101, 102, 201, 301, 401, 501, 601, 701, 801, 901, 902
```{r}
Plate = substr(df$Kernel_ID, start = 1, stop = 3)
keep_ind = Plate %in% c(101, 102, 201, 301, 401, 501, 601, 701, 801, 901, 902)
sum(keep_ind)

df_keep = df[keep_ind, ]
```

3. Create a class vector for aflatoxin. L means [AF] < 20 ppb, M means 20 ppb <= [AF] < 50 ppb, H means [AF] >= 50 ppb
```{r}
## Aflatoxin class
AF_class = ifelse(df_keep$AF.ppb. < 20, "L", "M")
AF_class[df_keep$AF.ppb. >= 50] = "H"
AF_class[is.na(df_keep$AF.ppb.)] = "L"
AF_class = factor(AF_class, levels = c("H", "M", "L"))
summary(AF_class)
```

4. Create a class vector for fumonisin. L means [FM] < 1 ppm, M means 1 ppm <= [FM] < 4 ppm, H means [FM] >= 4 ppm.
```{r}
## Fumonisin class
FM_class = ifelse(df_keep$FM.ppm. < 1, "L", "M")
FM_class[df_keep$FM.ppm. >= 4] = "H"
FM_class[is.na(df_keep$FM.ppm.)] = "L"
FM_class = factor(FM_class, levels = c("H", "M", "L"))
summary(FM_class)
```

5. Combine the data with the class vectors.
```{r}
Conc_lab = cbind(df_keep, AF_class, FM_class)
```

6. Output the file as Texas ELISA Conc and Label 04_12_2018.csv.
```{r, eval = FALSE}
write.csv(Conc_lab, "Texas ELISA Conc and Label 04_12_2018.csv")
```


### Visualization ###

1. Create a histogram for aflatoxin distribution grouped by class. Put the x axis on the log scale.
```{r}
AF_hist = ggplot(data = Conc_lab)+
  geom_histogram(aes(x = Conc_lab$AF.ppb., fill = Conc_lab$AF_class), bins = 50) +
  labs(x = "Aflatoxin Concentration (ppb)", y = "Number of Kernels")+
  scale_x_log10(breaks = c(1e-2, 1e-1, 1, 1e1, 1e2, 1e3, 1e4, 1e5)) +
  scale_fill_discrete(name = "Single Kernel Aflatoxin Level", labels = c("[AF]>=50", "[AF]<20")) +
  theme_classic() +
  theme(legend.position = "top")

AF_hist
```

2. Create a histogram for fumonisin distribution grouped by class. Put the x axis on the log scale.
```{r}
FM_hist = ggplot(data = Conc_lab)+
  geom_histogram(aes(x = Conc_lab$FM.ppm., fill = Conc_lab$FM_class), bins = 50) +
  labs(x = "Fumonisin Concentration (ppm)", y = "Number of Kernels")+
  scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1, 10, 100, 1000)) +
  scale_fill_discrete(name = "Single Kernel Fumonisin Level", labels = c("[FM]>=4", "1<=[FM]<4", "[FM]<1")) +
  theme_classic()+
  theme(legend.position = "top")

FM_hist
```

3. Create a bar chart for both aflatoxin and fumonisin grouped by class.
```{r, warning = FALSE}
barchart = ggplot(data = Conc_lab)+
  geom_bar(aes(x = rep("Aflatoxin", nrow(Conc_lab)), fill = AF_class), 
           position = "stack") +
  geom_bar(aes(x = rep("Fumonisin", nrow(Conc_lab)), fill = FM_class), 
           position = "stack") +
  geom_text(stat="count", aes(x = rep("Aflatoxin", nrow(Conc_lab)), 
                              fill = AF_class, 
                              label=..count..), 
            position = position_stack(vjust = 0.5)) +
  geom_text(stat="count", aes(x = rep("Fumonisin", nrow(Conc_lab)), 
                              fill = FM_class, 
                              label=..count..), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = "Number of Kernels") +
  scale_fill_discrete(name = "Single Kernel Mycotoxin Level", labels = c("High", "Low", "Medium")) +
  theme_classic()+
  theme(legend.position = "top")

barchart
```


4. Output the plots in a pdf file.
```{r, eval = FALSE}
pdf("Texas Corn Kernels ELISA Results.pdf")
  AF_hist
  FM_hist
  barchart
dev.off()
```

