---
title: "Texas Skewed Distribution 3"
author: "Xianbin Cheng"
date: "3/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Objective

  We want to create the weighted mycotoxin distribution figures with all of Texas 2017 corn data.
  
# Methods

1. Load the data and give a summary.

```{r, warning = FALSE, message = FALSE, echo = TRUE}
library(tidyverse)
library(kableExtra)
```

```{r, echo = TRUE}
# Mycotoxin concentrations
texas_conc = read.csv(file = "Texas_2017_combined_864obs.csv", header = TRUE, stringsAsFactors = FALSE, row.names = 1)
texas_conc$FM_class = ifelse(test = texas_conc$FM.ppm. >= 2, yes = "H", no = "L")
str(texas_conc)

# How many observations per class?
summary(as.factor(texas_conc$AF_class))
summary(as.factor(texas_conc$FM_class))

# Bulk mycotoxin levels
bulk_level = read.csv(file = "Bulk Level.csv", header = TRUE, stringsAsFactors = FALSE)
str(bulk_level)

# Weight and visible features
features = read.csv(file = "Texas_2017_weight_vis_features.csv", header = TRUE, stringsAsFactors = FALSE)
str(features)
```

2. Join the 3 datasets. First, `texas_conc` inner join `features` by `Kernel_ID` to get an intermediate data frame. Second, `bulk_level` right join that data frame. The final dataset should have 864 observations.

```{r, echo = TRUE}
full_data = inner_join(x = texas_conc, y = features, by = "Kernel_ID") %>%
  right_join(x = bulk_level, y = ., by = "Class")
colnames(full_data)[1] = "Bulk_class"
str(full_data)

# How many observations per bulk class?
summary(as.factor(full_data$Bulk_class))
```

3. Calculate how much mycotoxin each kernel contributes to the bulk level.
  
  **Two concepts:**
  
  * Prior bulk level = what Texas State Chemist provided
  * Posterior bulk level = weighted sum of mycotoxin concentrations = 
  
$$\frac {\sum_{i=1}^{n}C_im_i}{\sum_{i=1}^nm_i}$$ 
, where $C_i$ = mycotoxin concentration of kernel i, $m_i$ = mass of kernel i.

  Let's use the posterior bulk level for further analysis.
  
```{r, echo = TRUE}
a = full_data %>%
  group_by(Bulk_class) %>%
  mutate(wt_frac = m.kernel..g / sum(m.kernel..g), 
         wt_AF = AF.ppb. * wt_frac,
         wt_FM = FM.ppm. * wt_frac) 
```

4. Compare the prior bulk class level with the posterior bulk class level.

```{r, echo = TRUE}
b = a %>%
  group_by(Bulk_class) %>%
  summarise(Bulk_AF_prior = mean(Bulk_Aflatoxin.ppb.), Bulk_FM_prior = mean(Bulk_Fumonisin.ppm.),
            Bulk_AF_post = sum(wt_AF), Bulk_FM_post = sum(wt_FM))
```

# Result

```{r}
vis_wt_mycotoxin = function(data, type){

  if(type == "AF"){
    wt = "wt_AF"
    fill_class = "AF_class"
    fill_legend = "Single kernel aflatoxin level"
    ylab = "Weighted aflatoxin concentration (ppb)"
    
    temp = data %>%
      group_by(Bulk_class) %>%
      arrange(.data = ., desc(wt_AF)) %>%
      mutate(kernel_index = 1:n())
    
  } else if (type == "FM"){
    wt = "wt_FM"
    fill_class = "FM_class"
    fill_legend = "Single kernel fumonisin level"
    ylab = "Weighted fumonisin concentration (ppm)"
    
    temp = data %>%
      group_by(Bulk_class) %>%
      arrange(.data = ., desc(wt_FM)) %>%
      mutate(kernel_index = 1:n())
    
  } else {
    stop("Unknown mycotoxin type.")
  }
  
  ggplot(data = temp) +
    geom_col(aes_string(x = "kernel_index", y = wt, fill = fill_class)) +
    # scale_y_sqrt() +
    facet_wrap(~Bulk_class, scales = "free_y", ) +
    labs(x = "Kernel Index", y = ylab, fill = fill_legend) +
    scale_fill_discrete(label = c("High", "Low")) +
    theme_bw() +
    theme(legend.position = "top")
}
```

1. What are the prior and posterior bulk levels? Prior bulk levels are provided by the Texas State Chemist, and the posterior bulk levels are calculated by our experiments.

```{r, echo = FALSE}
kable_styling(kable_input = kable(x = b, format = "html", 
                                  col.names = c("Prior bulk class", "Prior bulk AF (ppb)", 
                                                "Prior bulk FM (ppm)", "Posterior bulk AF (ppb)", 
                                                "Posterior bulk FM (ppm)"), digits = 1), 
              full_width = TRUE)
```

2. Visualization for aflatoxin.

```{r}
vis_wt_mycotoxin(data = a, type = "AF")
```

3. Visualization for fumonisin

```{r}
vis_wt_mycotoxin(data = a, type = "FM")
```

# Appendix

```{r, echo = FALSE}
sessionInfo()
```

