---
title: "Texas Data Cleaning 6"
author: "Xianbin Cheng"
date: "March 5, 2019"
output: html_document
---

# Objective

  * Create a labeled dataset for Texas corn kernels collected in 2017
  
# Method

1. Load the libraries and files.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)
```

```{r}
sessionInfo()
```

```{r}
# Spec names and Kernel IDs
spec_kernel = read.csv("TX Spec_Kernel_ID_960obs.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = "")

# ELISA
Conc = read.csv("Texas_2017_combined_864obs.csv", header = TRUE, row.names = 1, stringsAsFactors = FALSE)

# Spectra
X1 = read.csv("TX X1.csv", header = TRUE, stringsAsFactors = FALSE)
X2 = read.csv("TX X2.csv", header = TRUE, stringsAsFactors = FALSE)
X1_norm = read.csv("TX X1_norm.csv", header = TRUE, stringsAsFactors = FALSE)
X2_norm = read.csv("TX X2_norm.csv", header = TRUE, stringsAsFactors = FALSE)
X1_perc = read.csv("TX X1_perc.csv", header = TRUE, stringsAsFactors = FALSE)
X2_perc = read.csv("TX X2_perc.csv", header = TRUE, stringsAsFactors = FALSE)
```

2. Combine `X1` with `X2` and combine `X1_norm` with `X2_norm`.

```{r}
X_combined = rbind(X1, X2)
X_norm_combined = rbind(X1_norm, X2_norm)
X_perc_combined = rbind(X1_perc, X2_perc)
```

3. Extract out the mycotoxin class labels. Join the `labels` with `spec_kernel` on `Kernel_ID`. Gather `Spec_ID1` and `Spec_ID2` into one `Name` colume. Remember, there are 48 NA's in `Spec_ID2` and we need to remove those.

```{r}
labels = Conc %>%
  dplyr::select(Kernel_ID, AF_class, FM_class)

temp = inner_join(x = labels, y = spec_kernel, by = "Kernel_ID") %>%
  gather(data = ., key = "Spec_type", value = "Spec_ID_all", Spec_ID1, Spec_ID2) %>%
  na.omit()

kernel_conc_spec = temp %>%
  dplyr::select(Kernel_ID, AF_class, FM_class, Spec_ID_all)
```

4. Join the `kernel_conc_spec` with `X_combined` and `X_norm_combined`. 

```{r}
X_Conc_1680_obs = inner_join(x = kernel_conc_spec, y = X_combined, by = c("Spec_ID_all" = "X"))
X_norm_Conc_1680_obs = inner_join(x = kernel_conc_spec, y = X_norm_combined, by = c("Spec_ID_all" = "X"))
X_perc_Conc_1680_obs = inner_join(x = kernel_conc_spec, y = X_perc_combined, by = c("Spec_ID_all" = "X"))
```

# Result

1. Summary of spectra before cleaning.

```{r}
str(spec_kernel)
str(Conc)

dim(X1)
dim(X2)
dim(X1_norm)
dim(X2_norm)
dim(X1_perc)
dim(X2_perc)
```

```{r}
# Spec_ID1: How many of them are not NA's?
spec_kernel %>%
  group_by(Bulk_Class) %>%
  dplyr::select( - Spec_ID2) %>%
  na.omit() %>%
  summarise(obs = n())

# Spec_ID2: How many of them are not NA's?
spec_kernel %>%
  group_by(Bulk_Class) %>%
  dplyr::select( - Spec_ID1) %>%
  na.omit() %>%
  summarise(obs = n())
```

2. Summary after cleaning.

```{r}
# How many spectra are there per bulk class?
temp %>%
  group_by(Bulk_Class) %>%
  summarise(obs = n())
```

```{r}
str(X_Conc_1680_obs, list.len = 10)
str(X_norm_Conc_1680_obs, list.len = 10)
str(X_perc_Conc_1680_obs, list.len = 10)
```

```{r, echo = FALSE, eval = FALSE}
write.csv(x = X_Conc_1680_obs, file = "TX X_Conc_1680_obs.csv")
write.csv(x = X_norm_Conc_1680_obs, file = "TX X_norm_Conc_1680_obs.csv")
write.csv(x = X_perc_Conc_1680_obs, file = "TX X_perc_Conc_1680_obs.csv")
```

